### 2016-07-27 mcmc.tex
renaming problems
boldface names
Adding problems for likelihood section

\begin{problem}
Run your M-H \MCMC\ sampler from \problemname~\ref{prob:MH}, but now
with a density function that is precisely unity \emph{everywhere}
(that is, at any input value of $x$ it returns unity).  That is, an
improper function (as discussed in \sectionname~\ref{sec:likelihood}).
Run it for longer and longer and plot the chain value $x$ as a
function of timestep.  What happens?
\end{problem}

\begin{problem}
For a real-world inference problem, read enough of
\citet{fittingaline} to understand and execute Exercise~6 in that
document.
\end{problem}

\begin{problem}
Modify the sampler you wrote in \problemname~\ref{prob:MH} to take
steps not in $x$ but in $\ln x$.  That is, replace the Gaussian
proposal distribution $q(x'\given x)$ with a Gaussian distribution in
$\ln x$ $q(\ln x'\given \ln x)$, but---incorrectly---make no other
changes.  What about your answers change?  How wrong is your answer?
Bonus points: Compute the analytic function that you have sampled from
(that is, not $p(x)$) and over-plot it on your histogram.
\end{problem}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/p8}
\end{center}
\caption{Solution to \problemname~\ref{prob:estimatetau}:
The autocorrelation function of the chain from \problemname~\ref{prob:MH}.}
\label{fig:estimatetau}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/p9}
\end{center}
\caption{Solution to \problemname~\ref{prob:itertau}:
Estimates of the integrated autocorrelation time of different segments of the
MCMC chain from \problemname~\ref{prob:MH} (black lines) and for the full
chain (orange line) as a function of window size $M$.
The ``optimized'' value computed using an iterative procedure is overplotted
as a dashed line and its value is listed in the title.}
\label{fig:itertau}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.48\textwidth]{problems/p10}
\includegraphics[width=0.48\textwidth]{problems/p11}
\end{center}
\caption{Solutions to \problemname s~\ref{prob:tuning} and~\ref{prob:tuningtau}:
Two methods of tuning the M--H proposal parameters.
\emph{left:} The acceptance fraction as a function of proposal scale for the
distribution from \problemname~\ref{prob:twod}a.
\emph{right:} The integrated autocorrelation time for each parameter
(indicated by the different colors) as a function of proposal scale parameter.}
\label{fig:tuning}
\end{figure}

>>>>

\begin{problem}\label{prob:improper}
Run your M-H \MCMC\ sampler from \problemname~\ref{prob:MH}, but now
with a density function that is precisely unity \emph{everywhere}
(that is, at any input value of $x$ it returns unity).  That is, an
improper function (as discussed in \sectionname~\ref{sec:likelihood}).
Run it for longer and longer and plot the chain value $x$ as a
function of timestep.  What happens?
\end{problem}

\begin{problem}\label{prob:fittingaline}
DFM: do this problem.
For a real-world inference problem, read enough of
\citet{fittingaline} to understand and execute Exercise~6 in that
document.
\end{problem}

\begin{problem}\label{prob:logprior}
Modify the sampler you wrote in \problemname~\ref{prob:MH} to take
steps not in $x$ but in $\ln x$.  That is, replace the Gaussian
proposal distribution $q(x'\given x)$ with a Gaussian distribution in
$\ln x$ $q(\ln x'\given \ln x)$, but make no other changes.
By doing this, you are no longer sampling the Gaussian $p(x)$ that you were in
\problemname~\ref{prob:MH}.
What about your answers change?  What distribution are you sampling now?
Compute the analytic function that you have sampled from~--~this will no longer
be the same $p(x)$~--~and over-plot it on your histogram.
\end{problem}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/logprior}
\end{center}
\caption{Solution to \problemname~\ref{prob:logprior}:
The black histogram shows the MCMC samples and the blue curve is the analytic
target density.}
\label{fig:estimatetau}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/estimatetau}
\end{center}
\caption{Solution to \problemname~\ref{prob:estimatetau}:
The autocorrelation function of the chain from \problemname~\ref{prob:MH}.}
\label{fig:estimatetau}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/itertau}
\end{center}
\caption{Solution to \problemname~\ref{prob:itertau}:
Estimates of the integrated autocorrelation time of different segments of the
MCMC chain from \problemname~\ref{prob:MH} (black lines) and for the full
chain (orange line) as a function of window size $M$.
The ``optimized'' value computed using an iterative procedure is overplotted
as a dashed line and its value is listed in the title.}
\label{fig:itertau}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.48\textwidth]{problems/tuning}
\includegraphics[width=0.48\textwidth]{problems/tuningtau}
\end{center}
\caption{Solutions to \problemname s~\ref{prob:tuning} and~\ref{prob:tuningtau}:
Two methods of tuning the M--H proposal parameters.
\emph{left:} The acceptance fraction as a function of proposal scale for the
distribution from \problemname~\ref{prob:twod}a.
\emph{right:} The integrated autocorrelation time for each parameter
(indicated by the different colors) as a function of proposal scale parameter.}
\label{fig:tuning}
\end{figure}
